Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/conclusions.tex
[ paper ] clausify added.

Summarizing, we provided for each \Metis inference rule a theorem in
type theory to emulate the rule reasoning. This approach, mainly
exposed in Section~\ref{ssec:emulating-inferences},
admits some improvements, we leave it for future work. We concern
mainly for a better coverage of the \simplify rule.
Other improvements are related with computational efficiency of
rules strongly depend on reordering tasks,
Section~\ref{sssec:reordering}.
We believe some rules like \canonicalize can get some help from an analysis done in the translator tool.

>>>>

Summarizing, we provided for each \Metis inference rule a theorem in
type theory to emulate the rule reasoning. This approach, mainly
exposed in Section~\ref{ssec:emulating-inferences},
admits some improvements that we leave it for future work.

For instance, we can investigate the consequences of removing $clausify$ inference by strengthen $canonicalize$ rule in Thm.~\ref{thm:canonicalize}
The \simplify rule is still lacking of some coverage since it is fairly complex, many cases could be omitted.
Other improvements are related with computational efficiency of
rules strongly depend on reordering tasks,
Section~\ref{sssec:reordering}.
We believe some rules like \canonicalize can get some help from an analysis done in the translator tool.

### 2017-10-22 paper/sections/proof-reconstruction/clausify.tex
[ paper ] roman in theorems.
[ paper ] table added.
[ paper ] clausify added.

The \clausify rule transforms a
propositional formula into its clausal normal form, a conjunction
of clauses. A \emph{clause} is the disjunction of zero or more
literals and a \emph{literal} is an atom (positive literal) or a
negation of an atom (negative literal).
It is important to notice that this kind of conversion
between one formula to its clausal normal form is not unique, and \Metis has a
customized approach to perform that transformation. That is the reason we
perform some reordering for the CNF of the source formula to fulfill this
issue.

In the following \Metis \TSTP derivation, we see an example where
\clausify expands the formula \texttt{normalize\_0\_0} by using distributive
laws presented in conjunctive normal form conversion.

\begin{verbatim}
fof(a1, axiom, (p => (q & r))).
...
fof(normalize_0_0, plain, (~ p | (q & r)),
    inference(canonicalize, [], [a1])).
fof(normalize_0_1, plain, ((~ p | q) & (~ p | r)),
    inference(clausify, [], [normalize_0_0])).
\end{verbatim}

\begin{definition}[clausify]
  \label{def:clausify}
 \begin{align*}
   \begin{split}
      &\fclausify : \Prop → \Prop → \Prop\\
      &\fclausify~φ~ψ =
      \begin{cases}
        ψ, &\text{ if }φ≡ψ\\
        \freorder_{∧∨}~(\fcnf~φ~ψ), &\text{ otherwise.}
      \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-clausify]
\label{thm:thm-clausify}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fclausify~φ~ψ$.
\end{theorem}

\begin{proof} If $φ ≡ ψ$, this lemmas follows after applying $\fsubst$
lemma. Otherwise, we use Theorem~\ref{thm:thm-reorder-and-or} and Theorem~\ref{thm:thm-cnf}.
\end{proof}

\begin{remark}
The \clausify rule is often preceded by the \canonicalize rule.
Both rules perform \emph{Clausification} that introduces
into the problem axioms or definitions in the domain of the solution.
The Clausification algorithms is mainly described on paper by
\citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}. We left for future work
investigate the consequences of removing $clausify$ inference by
strengthen $canonicalize$ rule in Def.~\ref{def:canonicalize}
\end{remark}

>>>>

The \clausify rule transforms a
propositional formula into its clausal normal form, a conjunction
of clauses. A \emph{clause} is the disjunction of zero or more
literals and a \emph{literal} is an atom (positive literal) or a
negation of an atom (negative literal).
It is important to notice that this kind of conversion between one formula 
to its clausal normal form is not unique, and \Metis has a customized 
approach to perform that transformation. That is the reason we perform some 
reordering for the \CNF of the source formula to fulfill this issue.

In the following \Metis \TSTP derivation, we see an example where
\clausify expands the formula \texttt{norm0} by using distributive
laws presented in conjunctive normal form conversion.

\begin{verbatim}
fof(a1, axiom, p => (q & r)).
...
fof(norm0, (~ p | (q & r)), inf(canonicalize, [a1])).
fof(norm1, ((~ p | q) & (~ p | r)), inf(clausify, [norm0])).
...
\end{verbatim}

\begin{mainth}
\label{thm:clausify}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fclausify~φ~ψ$, where,

  \begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fclausify : \Prop → \Prop → \Prop\\
  &\begin{array}{llll}
  \fclausify~φ~ψ &=
         \begin{cases}
        ψ, &\text{ if }φ≡ψ\\
        \freorder_{∧∨}~(\fcnf~φ~ψ), &\text{ otherwise.}
      \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}
If $φ ≡ ψ$, $Γ ⊢ \fclausify~φ~ψ$ normalizes to $Γ ⊢ ψ$. The conclusion follows by applying the $\fsubst$ lemma.\\[2mm]
Otherwise, we use Theorem~\ref{thm:thm-reorder-and-or} and Theorem~\ref{thm:thm-cnf}.
\end{proof}

\begin{remark}
The \clausify rule is often preceded by the \canonicalize rule.
Both rules perform \emph{Clausification} that introduces
into the problem axioms or definitions in the domain of the solution.
The clausification algorithms is mainly described on paper by
\citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}.
\end{remark}
